Tool-driven Revolutions in Archaeological Science

### 2018-09-09 analysis/paper/jcaa-paper.Rmd

The use of R or any other scripting languages enables a paradigm-shift, as it offers the possibility to share one's code. The sharing of code has been identified in other disciplines as an amplifier of methodological change [@touchon_and_mccoy_2016]. It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method. Thus, the sharing of code may be at the root of a tool-driven paradigm-change as desribed in the work of @Galison [-@Galison_1997]. This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@meltzer_1979, @trigger_2006]). Rather, this enables the archaologist to return to the tenet of the scientific method which demands the reproducibility of results. We do not want to claim archaeologists did not want to create reproducible research before the advance of scientific scripting languages. There is a longstanding tradition of publishing catalogues, lists and maps showing the data used for analyses. In the digital age, though, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses. This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results. Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which complicates the reproducibility of anlayses. As we could show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with biology (zooarchaeology) or focuses on scientific analyses.

Speaking with Kuhn's view on paradigm-change, a change in the view point of researchers seems to be happening in these areas: She or he, who works with computational methods, needs to view him/herself as a scientist, who needs to adhere to the scientific method, i. e. making the research reproducible for others.

We acknowledge, this is not an easy task. Learning a scripting language is time-intensive and writing code so that one is comfortable in sharing it even more so. [some more info from your case study maybe?]

@@

There is a lot of potential for other archaeologists to make their work reproducible, and we think this is a much needed paradigm-shift in computational archaeology. We therefore presented our freely available R package, rrtools, which is designed to make it easy for researchers in archaeology and other fields to make their research reproducible. We hope this will help to realise some of the potential to make archaeological research more open and reproducible and in this way encourage researchers to adhere to the scientific method.

>>>>

The use of R or any other scripting languages enables a paradigm-shift, as it offers the possibility to share one's code. The sharing of code has been identified in other disciplines as an amplifier of methodological change [@touchon_and_mccoy_2016]. It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method. Thus, the sharing of code may be at the root of a tool-driven paradigm-change as desribed in the work of @Galison [-@Galison_1997]. This change does not take place revolutionary-quick and does not replace older ideas as such (as has been a critique of Kuhn's idea-driven concept of paradigm shift by others [@meltzer_1979, @trigger_2006]). Rather, this enables the archaologist to return to the tenet of the scientific method which demands the reproducibility of results. We do not want to claim archaeologists did not want to create reproducible research before the advance of scientific scripting languages. There is a longstanding tradition of publishing catalogues, lists and maps showing the data used for analyses. In the digital age, though, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses. This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results. Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which complicates the reproducibility of analyses. As we could show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with biology (zooarchaeology) or focuses on scientific analyses.

We do not believe, that speaking with Kuhn's view on paradigm-change, a change in the view point of researchers seems to be happening: Data-and method sharing behaviour has been around in archaeology before. What we can see, though, is that the need for methods to share research in a way that is reproducible for others has been acknowledged and that archaeologists have looked for tools to make this possible. Using  scripting languages and data repositories is a solution, that enjoys increasing popularity.

We acknowledge, this is not an easy task. Learning a scripting language is time-intensive and writing code so that one is comfortable in sharing it even more so.

[some more info from your case study maybe?]

@@

There is a lot of potential for other archaeologists to make their work reproducible, and we think this is a much needed paradigm-shift in computational archaeology - linking the digital and computational times to  older practices of sharing data and methods. We therefore presented our freely available R package, rrtools, which is designed to make it easy for researchers in archaeology and other fields to make their research reproducible. We hope this will help to realise some of the potential to make computational archaeological research more open and reproducible and in this way encourage researchers to adhere to the scientific method.
